---
title: "Assignment 4 - Heart rate, respiration and interpersonal coordination"
author: "Matilde Jacobsen"
date: "November 7, 2019"
output: 
  md_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 4 - Heart rate, respiration and interpersonal coordination

Physiological data (here heart rate [variability], and respiration) are increasingly popular. Historically treated as pernicious noise to be regressed out of neuro-imaging data, there is now increasing research on how these signals tell us something important about cognition and beyond being just a signal of cognitive processes also impact them in interesting ways. Advanced sport science, and the quantified self movement (closely followed by marketing and communication) have hailed continuous physiological tracking as a powerful way to access and modify attitudes, habits, and performance. Further, as team coordination (in the military, in decision processes and organizational contexts) is more and more in focus, research has attempted to measure how interpersonal coordination between physiological systems might tell us something important about e.g. emotional and cognitive coordination. See references in the reading list for more on this.

In this assignment, you will learn to:
- collect physiological data
- pre-process physiological data (and grow further your mad R skills)
- model the continuous interdependence between two signals (using a multilevel model as proxy for a dynamical system approach)
- conservatively assess the presence of coordination between to signals in a controlled context

This assignment has two parts. The first part familiarizes you with heart rate, and respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.

2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

3) Do you observe differences in coordination between conditions? Report the models and results.

4) Is respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.

N.B. to give you a bit more data I included data from previous years (Study1, Study2 and Study 3). Note that synchronouns and turn-taking are the same across both studies, but the third condition is different: in the first year it was self-paced joint reading; in the second year it was the tv-series conversation.

## Let's get started

### Exploring physiological signals

- Choose one pair (one pair, three conditions)
- Load the logs
- Produce a plot of the participants' respiration signal and a different one of the participants' HR signal.
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3). There are also smarter packages, like cowplot and ggpubr.
- Can you eye-ball which condition if any displays more physiological coordination?

### First we read one data file and identify the procedure
- Load the file
- correctly identify all columns
- plot the data
- deal with the artifacts
- downsample the dat
- Add a column for study, group, trial and condition

```{r}
# Load the libraries
library(pacman, tidyverse)
p_load(ggplot2, gridExtra, tidyverse, groupdata2, purr, stats, lme4, lmerTest)

# Load the file
s1_t1 <- read.csv("Data/Study1_G1_T1_Synchronous.csv")
good <- read.csv("Data/Study3_G2_T2_Synchronous.csv")
# Plot
p1_resp <- ggplot(data = s1_t1) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")
p1_resp

p1_hr <- ggplot(data = s1_t1) +
  geom_path(aes(time, HR1, color = "P1")) +
  geom_path(aes(time, HR2, color = "P2")) +
  labs(x = "time", y = "HR") +
  theme(legend.position="bottom")
p1_hr
## Remove outliers

### Tip, check the function below
removeOuts <- function(ts,threshold){
  ts[ts > (mean(ts,na.rm=T) +
             (threshold*sd(ts,na.rm=T)))] = mean(ts,na.rm=T)+threshold*sd(ts,na.rm=T)
  ts[ts < (mean(ts,na.rm=T) -
             (threshold*sd(ts,na.rm=T)))] = mean(ts,na.rm=T)-threshold*sd(ts,na.rm=T)
  return(ts)
}
threshold=3 # Default value at 3 sds from the mean
#Removing outliers for respiration
s1_t1$Resp1_cleaned <- removeOuts(s1_t1$Resp1, threshold)
s1_t1$Resp2_cleaned <- removeOuts(s1_t1$Resp2, threshold)
#Removing outliers for HR
s1_t1$HR1_cleaned <- removeOuts(s1_t1$HR1, threshold)
s1_t1$HR2_cleaned <- removeOuts(s1_t1$HR2, threshold)

#Plotting for cleaned data
p1_resp_cleaned <- ggplot(data = s1_t1) +
  geom_path(aes(time, Resp1_cleaned, color = "P1")) +
  geom_path(aes(time, Resp2_cleaned, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")
p1_resp_cleaned

#Plotting for cleaned data
p1_hr_cleaned <- ggplot(data = s1_t1) +
  geom_path(aes(time, HR1_cleaned, color = "P1")) +
  geom_path(aes(time, HR2_cleaned, color = "P2")) +
  labs(x = "time", y = "HR") +
  theme(legend.position="bottom")
p1_hr_cleaned

# Plot raw data againt those with the artiacts removed
s1_t1_plots <- gridExtra::grid.arrange(p1_resp, p1_resp_cleaned,p1_hr,p1_hr_cleaned, nrow = 2, ncol = 2)

## Scale
### Tip: if scale() gives some issues, try the one below
z_scale <- function(column){
  column_c <- (column - mean(column)) / sd(column)
}
s1_t1_scaled <-s1_t1 %>% dplyr::mutate_at(c("Resp1_cleaned", "Resp2_cleaned", "HR1_cleaned", "HR2_cleaned"),z_scale)

# Plot again to check how scaled data look like
#Plotting for scaled data
p1_resp_scaled <- ggplot(data = s1_t1_scaled) +
  geom_path(aes(time, Resp1_cleaned, color = "P1")) +
  geom_path(aes(time, Resp2_cleaned, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")
p1_resp_scaled

#Plotting for scaled data
p1_hr_scaled <- ggplot(data = s1_t1_scaled) +
  geom_path(aes(time, HR1_cleaned, color = "P1")) +
  geom_path(aes(time, HR2_cleaned, color = "P2")) +
  labs(x = "time", y = "HR") +
  theme(legend.position="bottom")
p1_hr_scaled

s1_t1_plots_scaled <- gridExtra::grid.arrange(p1_resp_cleaned,p1_resp_scaled, p1_hr_cleaned, p1_hr_scaled, nrow = 2, ncol = 2)

## Downsample
### This is tricky, so you can have a look at my code  (relying on Ludvig's groupdata2) if you get stuck
s1_t1_scaled$row <- seq.int(nrow(s1_t1_scaled))
s1_t1_down = s1_t1_scaled %>%
  group(n = 100, method = 'greedy') %>%
  dplyr::summarise(
    time = mean(time,na.rm=T),
    HR1 = mean(HR1_cleaned,na.rm=T),
    HR2 = mean(HR2_cleaned,na.rm=T),
    Resp1 = mean(Resp1_cleaned,na.rm=T),
    Resp2 = mean(Resp2_cleaned,na.rm=T),
    rowname = row[1]) #the index we use to put them back together 

## Plot the downsampled data
p4 <- ggplot(data = s1_t1_down) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")
p4

## Now add the group, trial, condition to the cleaned up, scaled, downsampled data
match <- str_match('Study1_G1_T1_Synchronous.csv', "Study([\\d]+)_G([\\d]+)_T([\\d]+)_([:alpha:]+)")
x <- c("file", "study", "group", "trial", "condition")
colnames(match) <- x

s1_t1_down <- cbind(s1_t1_down, match)
## Tip the info is in the file name


```


## Now we are ready to go to load and pre-process all files

Go through all the files (with a function passed onto map_df), check which files should be excluded, if any, and save the pre-processed time-series

A couple of tips:
- looping is oh so slow. Making a function and using Map/Map_df is your salvation.
- each study restarts the group numbering, so you should make sure to change that (e.g. 100 * Study + Group)
- you need to make sure all the data are meaningful or something has to be removed. Plotting is your friend. E.g. "Study1_G1_T1_Synchronous" has one bad respiration signal. We could replace it with NAs


```{r}
## Remove outliers

### Tip, check the function below
removeOuts <- function(ts,threshold){
  ts[ts > (mean(ts,na.rm=T) +
             (threshold*sd(ts,na.rm=T)))] = mean(ts,na.rm=T)+threshold*sd(ts,na.rm=T)
  ts[ts < (mean(ts,na.rm=T) -
             (threshold*sd(ts,na.rm=T)))] = mean(ts,na.rm=T)-threshold*sd(ts,na.rm=T)
  return(ts)
}
threshold=3 # Default value at 2.5 sds from the mean

## Scale
### Tip: if scale() gives some issues, try the one below
z_scale <- function(column){
  column_c <- (column - mean(column)) / sd(column)
}

# Define a function running the loading, artifact removal, scaling, downsampling, info adding.

data_preprocess <- function(filename, threshold = 3){
  
  # To be filled in
  # load data
  file <- read.csv(filename)
  # change colnames
  colnames(file)[which(names(file) == "min")] <- "time"
  # list of HR and Respiration
  resp_hr <- c("Resp1", "Resp2", "HR1", "HR2")
  # change 
  file <- file %>% mutate_at(resp_hr, as.numeric)
  # remove outliers
  file$Resp1 <- removeOuts(file$Resp1, threshold)
  file$Resp2 <- removeOuts(file$Resp2, threshold)
  file$HR1 <- removeOuts(file$HR1, threshold)
  file$HR2 <- removeOuts(file$HR2, threshold)
  # scale
  file <- file %>% dplyr::mutate_at(resp_hr,z_scale)
  # downsample
  file$row <- seq.int(nrow(file))
  file <- file %>%
  group(n = 1000, method = 'greedy') %>%
  dplyr::summarise(
    time = mean(time,na.rm=T),
    HR1 = mean(HR1,na.rm=T),
    HR2 = mean(HR2,na.rm=T),
    Resp1 = mean(Resp1,na.rm=T),
    Resp2 = mean(Resp2,na.rm=T),
    rowname = row[1])
  # parse filename to extract study, diagnosis, subject and trial
  match <- str_match(filename, "Study([\\d]+)_G([\\d]+)_T([\\d]+)_([:alpha:]+)")
  x <- c("file", "study", "group", "trial", "condition")
  colnames(match) <- x
  # bind the matched and the downsampled data
  data <- cbind(match,file)
  # identifying and removing trials below a certain treshold
  dup_detector1 <- ifelse(duplicated(data$Resp1)==T,"duplicate","not-duplicate")
  dup_detector2 <- ifelse(duplicated(data$Resp2)==T,"duplicate","not-duplicate")
   if(sum(dup_detector1=="duplicate") >= 8 | sum(dup_detector2=="duplicate") >= 8) {
    file$Resp1 <- NA
    file$Resp2 <- NA
   } 
     
  return(data)

}

# test it on just one file while writing the function
test_data <- data_preprocess("Data/Study1_G2_T1_TurnTaking.csv")
test_read <- data_preprocess("Data/Study1_G1_T1_Synchronous.csv")

#  Identify all files to be read
# Run the function on the whole dataset using map_df
phys_data <- list.files(path = "/Users/matilde/Desktop/AU/Experimental Methods III/Assignment 4 - heartrate_respiration/assignment-4/data",pattern = ".csv", full.names = T) %>% ## NB replace with your path to the files
    purrr::map_df(data_preprocess)
```

After preprocess we need to assess the data and change some variables
```{r}
phys_data_up <- phys_data
# Making a unique pair ID
phys_data$study <- as.numeric(phys_data$study)
phys_data$group <- as.numeric(phys_data$group)

phys_data$uPairID <- (100 * phys_data$study + phys_data$group)

#Normalizing the time values
#Assuming that the large values are millisecond
#choose 400 arbitrarily because it is above a reasonable minute count.
phys_data[which(phys_data$time > 400),]$time <- phys_data[which(phys_data$time > 400),]$time / 1000 / 60
 
#time since 0
phys_data <- phys_data %>% group_by(uPairID, trial) %>% mutate(actual_time_min = time - min(time))

# change timename
colnames(phys_data)[7] <- "time_min"

# Argumentation for why we use 100 as threashold
overview <- phys_data_up %>% 
  group_by(file) %>% 
  summarise(
    sum(duplicated(Resp1)),
    sum(duplicated(Resp2)))

overview_down <- phys_data %>% 
  group_by(file) %>% 
  summarise(
    sum(duplicated(Resp1)),
    sum(duplicated(Resp2)))

# Now we need to make sure all the data are meaningful or something has to be removed
# E.g. "Study1_G1_T1_Synchronous" has one bad respiration signal. We could replace it with NAs
summary(phys_data$condition)
# plots plots plots
phys_data %>% subset(study==4) %>%
  ggplot() + 
  geom_line(aes(actual_time_min,HR1),color="red") + 
  geom_line(aes(actual_time_min,HR2),color="blue") + 
  facet_grid(group ~ trial)

phys_data %>% subset(study==4) %>%
  ggplot() + 
  geom_line(aes(actual_time_min,Resp1),color="red") + 
  geom_line(aes(actual_time_min,Resp2),color="blue") + 
  facet_grid(group ~ trial)

phys_data_up %>% subset(study==4) %>%
  ggplot() + 
  geom_line(aes(actual_time_min,Resp1),color="red") + 
  geom_line(aes(actual_time_min,Resp2),color="blue") + 
  facet_grid(group ~ trial)

# Romove bad heart rate data
# Removing after 3 min.
phys_data$HR1 <- ifelse(phys_data$actual_time_min > 3 & phys_data$study == 3, NA, phys_data$HR1)
phys_data$HR2 <- ifelse(phys_data$actual_time_min > 3 & phys_data$study == 3, NA, phys_data$HR2)
phys_data$Resp1 <- ifelse(phys_data$actual_time_min > 3 & phys_data$study == 3, NA, phys_data$Resp1)
phys_data$Resp2 <- ifelse(phys_data$actual_time_min > 3 & phys_data$study == 3, NA, phys_data$Resp2)

#Removing single bad trials
#Study 2: G5T1-3, G6T1-3, G8T1-3, G9T1
#Study 3: G9T1-3, G1T3        
#Study 4: G3T1-3, G5T1-3, G6T1-3, G7T1-3, G8T
bad_files <- c("Study2_G5_T1_Synchronous","Study2_G5_T2_TurnTaking","Study2_G5_T3_Conversation","Study2_G6_T1_TurnTaking","Study2_G6_T2_Conversation","Study2_G6_T3_Synchronous","Study2_G8_T1_TurnTaking","Study2_G8_T2_Synchronous","Study2_G8_T3_Conversation","Study2_G9_T1_Synchronous","Study3_G9_T1_Conversation","Study3_G9_T2_Synchronous","Study3_G9_T3_TurnTaking","Study3_G1_T3_Conversation","Study4_G3_T1_MovementGuided","Study4_G3_T2_MovementCoop","Study4_G3_T3_Synchronous","Study4_G5_T1_Synchronous","Study4_G5_T2_TurnTaking","Study4_G5_T3_Conversation","Study4_G6_T3_Conversation","Study4_G6_T1_TurnTaking","Study4_G6_T2_Synchronous","Study4_G7_T1_MovementGuided","Study4_G7_T2_MovementCoop","Study4_G7_T3_Synchronous","Study4_G8_T4_MovementCoop")

phys_data$HR1 <- ifelse(phys_data$file %in% bad_files , NA, phys_data$HR1)
phys_data$HR2 <- ifelse(phys_data$file %in% bad_files , NA, phys_data$HR2)



# Save the data
write.csv(phys_data, file = "physiological_data.csv")
phys_data <- read.csv("physiological_data.csv")    


```

Plotting individual trials
```{r}
bad1 <- data_preprocess("Data/Study2_G7_T3_TurnTaking.csv")
bad2 <- data_preprocess("Data/Study1_G1_T2_TurnTaking.csv")
bad3 <- data_preprocess("Data/Study1_G1_T1_Synchronous.csv")
bad4 <- data_preprocess("Data/Study3_G9_T2_Synchronous.csv")
bad5 <- data_preprocess("Data/Study4_G4_T5_TurnTaking.csv")
bad6 <- data_preprocess("Data/Study4_G1_T1_Synchronous.csv")
bad7 <- data_preprocess("Data/Study4_G4_T1_MovementGuided.csv")
good <- data_preprocess("Data/Study3_G2_T2_Synchronous.csv")
# Plot
p1 <- ggplot(data = bad1) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p2 <- ggplot(data = bad2) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p3 <- ggplot(data = bad3) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p4 <- ggplot(data = bad4) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p5 <- ggplot(data = bad5) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p6 <- ggplot(data = bad6) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p7 <- ggplot(data = bad7) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")

p_good <- ggplot(data = good) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time", y = "Resp") +
  theme(legend.position="bottom")
p_good
plot_collection <- gridExtra::grid.arrange(p2, p3, p4, p5, p6, p7)
gridExtra::grid.arrange(p2,p_good)
grid.arrange(arrangeGrob(p2, top="Bad plot"), arrangeGrob(p_good, top="Good plot"))


```

## Now we need to run some analysis

Let's start with a multilevel model that accounts for 
- stability (how each signal is autocorrelated)
- interpersonal dependence (each signal is dependent from the previous state of the other signal)

The data needs to be further prepared, so we can analyze both participants in the same model.
We need to turn the data into a long format:
- a column indicating own hr and one own respiration
- a column indicating other hr and one other respiration
- a column indicating change in hr from previous round and one in respiration

We can then run an analysis where change is a function of one's previous state (stability, see slides), and the other's previous state (coupling). Make sure to:
- set up the most interesting contrasts: how do these parameters vary by condition? which condition should be baseline?
- set up the right random effects.
- N.B. the model will be slow. Make sure it works on a subset of the data first!

Bonus question: what if we include an additional layer? Is my heart rate just adjusting to yours, or also to how much you are adjusting to mine?
- to start answering this we can add a column indicating the previous change in hr in the other and one in respiration
- we can then build on the previous models by also adding the previous change in the other


```{r}
phys_data <- read.csv("physiological_data.csv")
#lme4 can do a trick to invert betas by adding - 1  
# Genearate a column for each: previous HR1, HR2, Resp1, Resp2
phys_data$study <- as.numeric(phys_data$study)
phys_data$group <- as.numeric(phys_data$group)
phys_data$trial <- as.numeric(phys_data$trial)

phys_data <- phys_data %>% 
  group_by(group, study, trial) %>%
  mutate(HR1_future = lead(HR1, 1))

phys_data <- phys_data %>% 
  group_by(group, study, trial) %>%
  mutate(HR2_future = lead(HR2, 1))

phys_data <- phys_data %>% 
  group_by(group, study, trial) %>%
  mutate(Resp1_future = lead(Resp1, 1))

phys_data <- phys_data %>% 
  group_by(group, study, trial) %>%
  mutate(Resp2_future = lead(Resp2, 1))

# Genearate a column for each: change in HR1, HR2, Resp1, Resp2
## create variable for future HR1 and HR2 (HR1_future, HR2_future), using lag(x-1), group by study/trial/condition, use mutate to put it in the same function

phys_data <- phys_data %>% 
  mutate(
    HR1_change = (HR1_future - HR1),
    HR2_change = (HR2_future - HR2),
    Resp1_change = (Resp1_future - Resp1),
    Resp2_change = (Resp2_future - Resp1)
  )

# Make the data long, so we can analyze both participants at the same time 
## N.B. This is a bit tricky and you might have to do it in several steps
d_hr_future <- 
  gather(phys_data, # data        
         participant, HR_future, # new vars
         HR1_future, HR2_future) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, HR_future, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

d_resp_future <- 
  gather(phys_data, # data        
         participant, Resp_future, # new vars
         Resp1_future, Resp2_future) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, Resp_future, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

d_hr_change_self <- 
  gather(phys_data, # data        
         participant, HR_change_self, # new vars
         HR1_change, HR2_change) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, HR_change_self, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

d_resp_change_self <- 
  gather(phys_data, # data        
         participant, Resp_change_self, # new vars
         Resp1_change, Resp2_change) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, Resp_change_self, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

d_hr_self <- 
  gather(phys_data, # data        
         participant, HR_self, # new vars
         HR1, HR2) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, HR_self, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

d_resp_self <- 
  gather(phys_data, # data        
         participant, Resp_self, # new vars
         Resp1, Resp2) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, Resp_self, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

phys_data_other <- phys_data %>% 
  rename(
    HR1 = "HR2",
    HR2 = "HR1",
    Resp1 = "Resp2",
    Resp2 = "Resp1",
    HR1_change = "HR2_change",
    HR2_change = "HR1_change",
    Resp1_change = "Resp2_change",
    Resp2_change = "Resp1_change"
  )

d_hr_change_other <- 
  gather(phys_data_other, # data        
         participant, HR_change_other, # new vars
         HR1_change, HR2_change) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, HR_change_other, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

d_resp_change_other <- 
  gather(phys_data_other, # data        
         participant, Resp_change_other, # new vars
         Resp1_change, Resp2_change) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, Resp_change_other, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

d_hr_other <- 
  gather(phys_data_other, # data        
         participant, HR_other, # new vars
         HR1, HR2) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, HR_other, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

d_resp_other <- 
  gather(phys_data_other, # data        
         participant, Resp_other, # new vars
         Resp1, Resp2) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, Resp_other, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

dd <- cbind(d_hr_change_self,d_hr_change_other,d_resp_change_self,d_resp_change_other,d_hr_future,d_resp_future,d_hr_self,d_hr_other,d_resp_self,d_resp_other) 
dd <- dd %>%  select(actual_time_min,participant,group,condition,trial,study,HR_self,Resp_self,HR_other,Resp_other,HR_future,Resp_future,HR_change_self,HR_change_other,Resp_change_self,Resp_change_other)
#Making participant unique for study
dd$participant <- (dd$study * 1000)+dd$participant

# Set the most interesting contrast e.g. by defining synchronous or conversation as the baseline
# We only want to analyse study 4 - making a subset of the data
dd4 <- subset(dd,study == 4)
# Releveling with conversation as baseline
dd4$condition <- as.factor(dd4$condition)
dd4$condition <- relevel(dd4$condition, ref = "Conversation")

ggplot(dd_single, aes(HR_self,HR_change_self)) +
  geom_smooth(aes(HR_self, HR_other, method = "lm"))
# effects
# Plotting for single participant
ggplot(data = dd_single, aes(x = HR_other, y = HR_change_self)) + geom_smooth(aes(HR_self, HR_other, method = "lm"))
ggplot(data = dd_single) +
  geom_point(aes(HR_self, HR_change_self, color = "HR_self")) +
  geom_point(aes(HR_other, HR_change_self, color = "HR_other")) +
  geom_smooth(aes(HR_self, HR_change_self, method = "lm", color = "HR_self")) +
  geom_smooth(aes(HR_other, HR_change_self, method = "lm", color = "HR_other")) +
  labs(x = "HR", y = "HR_change") +
  theme(legend.position="bottom")  
              
# Model change as a function of own and other previous state 
  #stability = beta HR_self
  #coupling = beta HR_other
  #baseline = surrogate pairs

#Subset single participant 101
dd_single <- subset(dd4, participant == 4101)

#Single participant
single_model1 <- lm(HR_change_self ~ 1 + HR_self + HR_other, data = dd_single)
single_model1
#Subset single group and run model
dd_group1 <- subset(dd4, group == 1)
model_group1 <- lmer(HR_change_self ~ 0 + condition + (HR_self + HR_other) : condition + 
                          (0 + condition | participant), data = dd_group1)

#Model for study 4 only
model_study4 <- lmer(HR_change_self ~ 0 + condition + (HR_self + HR_other) : condition + 
                          #we tell the model that we are not all the same. Accouting for independance of the                           data. Plus we are doing partial pooling.
                          (0 + condition | group) +        
                          (0 + condition | participant), data = dd4)
#Model from interaction effect
model_study4_interaction <- lmer(HR_change_self ~ 1 + condition + (HR_self + HR_other) * condition + 
                          #we tell the model that we are not all the same. Accouting for independance of the                           data. Plus we are doing partial pooling.
                          (1 + condition | group) +        
                          (1 + condition | participant), data = dd4)
summary(model_study4)
summary(model_study4_interaction)
#alternative main effect instead of HR_other = HR_other-self

# Bonus points: Add to the previous model also change in the other to see whether my adaptation is influenced by the other's adaptation.

```


## Now we need to create control baselines.

First shuffled controls, then surrogate pairs.

### Creating controls: shuffled controls

Shuffled controls break the temporal dependencies of time-series by shuffling the value within one time-series. This ensures the "coordination" observed is not due to the actual values in the series and not their sequence.
Tip: sample() is your friend, but make sure to shuffle things within participant/condition and not throughout the whole dataset
 
```{r}

# Create a shuffled dataset

# Concatenate it to the original dataset (and remember to have a column telling you which is which)

# Create the same models as in the previous chunk, but adding an interaction by shuffled vs. real
#The baseline by condition might vary by type
model_control <- lmer(HR_change_self ~ 0 + (condition + (HR_self + HR_other) : condition) : type #surrogate vs. real + 
                          (0 + condition | participant), data = dd)
```
 
 
### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair)

```{r}

# Identify unique pairs within a given study (to keep things manageable) and create list of possible surrogate pairs (e.g. individual 1 from pair 1 and individual 2 from pair 2)

# Starting from the wide format, create "surrogate" dataset with the data from surrogate pairs
groups <- as.numeric(as.character(unique(dd$group[dd$study==4]))) # List all pairs
SurrogateList <- expand.grid(a = groups, b = groups) # Identify all possible combinations of 2 pairs
SurrogateList = subset(SurrogateList, a != b)  # exclude combinations with identical pairs

phys_data4 <- subset(phys_data, study == 4)
#Create new dataframe
surrogate_data <- phys_data4[0,]

for (i in 1:nrow(SurrogateList)){  # loop through all combinations
  x <- subset(phys_data4, group==SurrogateList$a[i]) # subset data from the first pair    
  y <- subset(phys_data4, group!=SurrogateList$a[i]) # subset data from the second pair   
  newPairID <- c(800 + ((1:4)*i))                            
  # create new pair id
    for (co in c("Synchronous","TurnTaking","Conversation", "MovementCoop","MovementGuided")){ # loop through conditions
    if (co %in% unique(x$condition) & co %in% unique(y$condition)){ # check that both pairs have the data for that condition
      z1 <- subset(x, condition==co) # subset only that condtion from first pair
      z2 <- subset(y, condition==co) # subset only that condtion from second pair
      if (nrow(z1) > nrow(z2)) {	# make sure data have same length in both pairs
        z1<-z1[1:nrow(z2),]
        }
      if (nrow(z2) > nrow(z1)) { 
        z2<-z2[1:nrow(z1),]
        }
      w1 <- z1 %>% mutate(	# assemble new pair combining the 2 pairs
        HR2 = z2$HR2,
        Resp2 = z2$Resp2,
        HR2_future = z2$HR2_future, 
        Resp2_future = z2$Resp2_future, 
        HR2_change = z2$HR2_change, 
        Resp2_change = z2$Resp2_change)
      # Saving data
      if (nrow(surrogate_data) == 0) {
        surrogate_data <- w1
        }
        else {
          surrogate_data <- rbind(surrogate_data,w1)
        } # make sure that you do this!
}}}
# Make it into long format
hr_future <- 
  gather(surrogate_data, # data        
         participant, HR_future, # new vars
         HR1_future, HR2_future) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, HR_future, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

resp_future <- 
  gather(surrogate_data, # data        
         participant, Resp_future, # new vars
         Resp1_future, Resp2_future) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, Resp_future, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

hr_change_self <- 
  gather(surrogate_data, # data        
         participant, HR_change_self, # new vars
         HR1_change, HR2_change) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, HR_change_self, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

resp_change_self <- 
  gather(surrogate_data, # data        
         participant, Resp_change_self, # new vars
         Resp1_change, Resp2_change) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, Resp_change_self, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

hr_self <- 
  gather(surrogate_data, # data        
         participant, HR_self, # new vars
         HR1, HR2) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, HR_self, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

resp_self <- 
  gather(surrogate_data, # data        
         participant, Resp_self, # new vars
         Resp1, Resp2) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, Resp_self, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

surrogate_data_other <- surrogate_data %>% 
  rename(
    HR1 = "HR2",
    HR2 = "HR1",
    Resp1 = "Resp2",
    Resp2 = "Resp1",
    HR1_change = "HR2_change",
    HR2_change = "HR1_change",
    Resp1_change = "Resp2_change",
    Resp2_change = "Resp1_change"
  )

hr_change_other <- 
  gather(surrogate_data_other, # data        
         participant, HR_change_other, # new vars
         HR1_change, HR2_change) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, HR_change_other, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

resp_change_other <- 
  gather(surrogate_data_other, # data        
         participant, Resp_change_other, # new vars
         Resp1_change, Resp2_change) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, Resp_change_other, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

hr_other <- 
  gather(surrogate_data_other, # data        
         participant, HR_other, # new vars
         HR1, HR2) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, HR_other, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

resp_other <- 
  gather(surrogate_data_other, # data        
         participant, Resp_other, # new vars
         Resp1, Resp2) %>% #old vars
  select( # drop irrelevant vars
    actual_time_min, Resp_other, participant, study, group, condition, trial) %>%
  mutate( # create unique participant ID
    participant = parse_number(as.character(group)) * 100 + 
      parse_number(participant))

dd2 <- cbind(hr_change_self,hr_change_other,resp_change_self,resp_change_other,hr_future,resp_future,hr_self,hr_other,resp_self,resp_other) 
dd2 <- dd2 %>%  select(actual_time_min,participant,group,condition,trial,study,HR_self,Resp_self,HR_other,Resp_other,HR_future,Resp_future,HR_change_self,HR_change_other,Resp_change_self,Resp_change_other)

#Making participant unique for study
dd2$participant <- (dd2$study * 1000)+dd2$participant

#Create type variable
dd2$type <- "surrogate"
dd$type <- "real"

#Bind the dataframes together
dd_all <- rbind(dd,dd2)
#Creating new main effect for other - self
dd_all$other_self <- (dd_all$HR_other) - (dd_all$HR_self)
#Subset study 4
dd_all_4 <- subset(dd_all, study == 4)
# Create models as in chunks above, but adding an interaction with the Real vs. Surrogate variable (exclude shuffled ones for simplicity)
model_surrogate <- lmer(HR_change_self ~ 0 + (condition + (HR_self + HR_other):condition):type + 
                          (1 + condition | group) +        
                          (1 + condition | participant), 
                        data = dd_all_4, 
                        REML = F, 
                        control = lmerControl(
                        optimizer = "nloptwrap",
                        calc.derivs = FALSE,
                        optCtrl=list(xtol_abs=1e-8, ftol_abs=1e-8, maxfun = 1000)))
summary(model_surrogate)
# Model with other-self as main effect
model_control2 <- lmer(HR_change_self ~ 0 + (condition + (HR_self + other_self):condition):type + 
                          (1 + condition | group) +        
                          (1 + condition | participant), 
                        data = dd_all_4, 
                        REML = F, 
                        control = lmerControl(
                        optimizer = "nloptwrap",
                        calc.derivs = FALSE,
                        optCtrl=list(xtol_abs=1e-8, ftol_abs=1e-8, maxfun = 1000)))

summary(model_control2)

# Releveling with conversation as baseline
dd_all_4$type <- as.factor(dd_all_4$type)
dd_all_4$type <- relevel(dd_all_4$type, ref = "surrogate")

```
 

### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them

 